### 2014-03-02 paper/formal.tex
pass over section 3

In this section we present the formal security guarantees provided by
our embedding.
%
Concretely, we show that |specLang alpha targetLang| for an appropriate
scheduler, satisfies non-interference~\tocite{},
without making any reference to properties of |targetLang|.
%
As it turns out, the precise scheduling policy |alpha| dictates
whether or not the guarantee holds for programs with diverging tasks.
For a sequential scheduler, we will only be able to show
\emph{termination-insensitive non-interference}, where a program
may diverge based on secret data; this allows attackers to observe
secrets by observing the termination of tasks.
For the concurrent round-robin schedule,
we can show a stronger result known as
\emph{termination-sensitive non-interference},
where termination attacks cannot leak information.

How can we characterize what the \emph{secret inputs} of a program
are?  Like other work,~\tocite{} we specify and prove
non-interference using \emph{term erasure}.
%
Intuitively, in this approach we proving that an attacker does not learn
any sensitive information from a program by 
Intuitively, term erasure allows us to show that a an
attacker does not learn any sensitive information from a program if the
program behaves identically to a program with all sensitive data ``erased''.
%
To accommodate for this, we extend our formalism to account for
programs with erased configurations and a function |erasef il| that
performs erasures, usually by rewriting (parts of) configurations that
are more sensitive than |il| to a new syntactic construct |bullet|.
%
We can characterize the types of languages for which one can state
a non-interference claim.

\begin{definition}[Information flow control language]
    An information flow control language |L| is a tuple |(C, .->,
    erasef l)|, where $|C|$ is the type of machine configurations (members
    of which are usually denoted by the meta-variable |c|), |.->| is a
    reduction relation between machine configurations and |erasef l : C -> tyerase C|
    is an erasure function parametrized on labels from machine configurations to \emph{erased} machine
    configurations |tyerase C|.  This tuple is sometimes
    augmented with a set of terminal configuration states $V
    \subseteq |C|$ for which there are no possible further transitions.
\end{definition}

@@

Non-interference relies on the definition of an attacker's observational power at
security level |l|, which is typically defined as an equivalence
relation---called |l|-equivalence---on configurations.  An attacker
at level |l| cannot
distinguish two configurations that are |l|-equivalent.
%
Since our erasure function captures the attacker model, we simply define this
equivalence as the syntactic-equivalence of erased configurations.
%
\begin{definition}[|l|-equivalence]
    In a language |(C, .->, erasef l)|, two machine configurations
    |memberf (c, c') C| are considered $l$-equivalent, written as |c ~= c'|,
    if |erase l c = erase il c'|.
\end{definition}
%

@@

\begin{definition}[Termination Sensitive Non-Interference (TSNI)]
  A language |(C, .->, erasef l)| is termination
  sensitive non-interfering if for any label |l|, and configurations
  |c1, c1', c2 member C|, if
  \begin{equation} \label{eq:tsni-lhs}
    |c1| \approx_{|l|} |c2|
    \qquad \text{and} \qquad
    |c1| |.->|^* |c1'|
  \end{equation}
  then there exists a configuration |c2' member C| such that
  \begin{equation} \label{eq:tsni-rhs}
    |c1'| \approx_{|l|} |c2'|
     \qquad \text{and} \qquad
    |c2| |.->|^* |c2'|
    \ \text{.}
  \end{equation}
\end{definition}
%
In other words if we take two |l|-equivalent configurations, then for every
intermediate step taken by the first configuration, there is a corresponding
number of steps that the second configuration can take to result in a
configuration that is |l|-equivalent to the first resultant configuration.
\Red{TODO: Do we want a commutative diagram to illustrate this?}

We now show that our IFC language satisfies TSNI under a particular
scheduling policy.
%
Specifically, we consider the language under the round-robin scheduler
|roundrobinf| given in Figure~\ref{fig:scheduler}.

@@

In general TSNI will not hold for an arbitrary scheduler |alpha|.
%
For example, |specLang alpha targetLang| with a scheduler that inspects a
sensitive threads current state when deciding which thread to schedule next
will in general not satisfy TSNI.
%
%In Section~\toref{} we detail the preceise scheduler for which our
%specification language is TSNI.
%

However, even non-adversairal schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI.
%
Concretely, consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |pub canFlowTo sec|, but |sec cantFlowTo pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(when false diverge)),  fullconf 2 pub tS1 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(when true diverge)),   fullconf 2 pub tS2 ie
\end{code}
These two configurations are |pub|-equivalent, but |ic1| will reduce (in two
steps) to |ic1 = iS ; fullconf 2 pub tS1 (IT diverge)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Setion~\ref{sec:extensions}.}
then the |sec| tasks's decision to diverge or not is directly leaked to a
public entitiy.

% sheule: I feel like this forward reference is not useful here.
%In certain cases, however, a TSNI-safe concurrent scheduling policy such as
%|roundrobinf| cannot be considered without imposing a non-trivial burden on the
%concrete implementation.
%(Section~\ref{ref:real} describes a particular embedding with JavsScript as the
%target language.)
%
To accomodate for sequential languages, or cases where a weaker guarantee
is sufficient, we consider an alternative non-interference property called termination insensitive
non-interference (TINI).  This property can also be upheld by sequential languages at the cost
of leaking one bit through (non)-termination.
%
\begin{definition}[Termination insensitive non-interference (TINI)]
  A language |(C, V, .->, erasef l)| is termination
  insensitive non-interfering if for any label |l|, and configurations
  |c1, c2 member C| and |c1', c2' member V|, it holds that
  \[
    (|c1| \approx_{|l|} |c2|
    \;\;\land\;\;
    |c1| |.->|^* |c1'|
    \;\;\land\;\;
    |c2| |.->|^* |c2'|)
    \implies
    |c1'| \approx_{|l|} |c2'|
  \]
\end{definition}

@@

Nevertheless, we believe TINI to be a useful for many scenarios and thus we
show that our IFC language with the sequential scheduling policy |seqf|
satisfies TINI.
%
\begin{theorem}[Sequential IFC language is TINI]
  \label{thm:seq-tini}
For any target language |targetLang|, |specLang seqf targetLang| satisfies TINI.
\end{theorem}

>>>>

In this section we present the formal security guarantees provided by
our embedding.
%
Concretely, we show that |specLang alpha targetLang| for, appropriate
scheduler, satisfies non-interference~\cite{Goguen82},
without making any reference to properties of |targetLang|.
%
%As it turns out, 
The precise scheduling policy |alpha| dictates
the guarantee holding for programs with diverging tasks.
For a sequential scheduler |seqf|, we will only be able to show
\emph{termination-insensitive non-interference}, where a program
may diverge based on secret data; this allows attackers to observe
secrets by observing the termination of tasks.
For the concurrent round-robin schedule |roundrobinf|,
we can show a stronger result known as
\emph{termination-sensitive non-interference},
where termination attacks cannot leak information.

How can we characterize what the \emph{secret inputs} of a program are?  Like
other work~\cite{Li+:2010:arrows,Russo+:Haskell08,lio,stefan:addressing-covert},
we specify and prove non-interference using \emph{term erasure}.
%
%Intuitively, we prove that an attacker does not learn
%any sensitive information from a program by 
Intuitively, term erasure allows us to show that a an attacker does not learn
any sensitive information from a program if the program behaves identically
(from the attacker point of view) to a program with all sensitive data
``erased''.
%
To accommodate for this, we extend our formalism to account for
programs with erased configurations and a function |erasef il| that
performs erasures, usually by rewriting (parts of) configurations that
are more sensitive than |il| to a new syntactic construct |bullet|.
%
We can characterize the types of languages for which one can state
a non-interference claim.

\Red{should be the environment colored in the definition?}
\begin{definition}[Information flow control language]
    An information flow control language |L| is a tuple |(C, .->,
    erasef l)|, where $|C|$ is the type of machine configurations (members
    of which are usually denoted by the meta-variable |c|), |.->| is a
    reduction relation between machine configurations and |erasef l : C -> tyerase C|
    is an erasure function parametrized on labels from machine configurations to \emph{erased} machine
    configurations |tyerase C|.  This tuple is sometimes
    augmented with a set of terminal configuration states $V
    \subseteq |C|$ for which there are no possible further transitions.
    \Red{It is a bit odd to extend a tuple ``sometimes'' and a |V| appears like from
      nowhere. What about saying: Sometimes, we refer as |V| to set of 
     terminal configurations in |C|, i.e., configurations states where
     no further transitions are possible. If you agree, then the next sentence
     where you said |V = ...| must be removed.}
\end{definition}

@@

Non-interference relies on the definition of an attacker's observational power at
security level |l|, which is typically defined as an equivalence
relation---called |l|-equivalence---on configurations.  An attacker
at level |l| cannot
distinguish two configurations that are |l|-equivalent.
%
Since our erasure function captures the attacker model, we simply define this
equivalence as the syntactic-equivalence of erased configurations~\cite{stefan:addressing-covert}.
%
\begin{definition}[|l|-equivalence]
    In a language |(C, .->, erasef l)|, two machine configurations
    |memberf (c, c') C| are considered $l$-equivalent, written as |c ~= c'|,
    if |erase l c = erase il c'|.
\end{definition}
%

@@

\begin{definition}[Termination Sensitive Non-Interference (TSNI)]
  A language |(C, .->, erasef l)| is termination
  sensitive non-interfering if for any label |l|, and configurations
  |c1, c1', c2 member C|, if
  \begin{equation} \label{eq:tsni-lhs}
    |c1| \approx_{|l|} |c2|
    \qquad \text{and} \qquad
    |c1| |.->|^* |c1'|
  \end{equation}
  then there exists a configuration |c2' member C| such that
  \begin{equation} \label{eq:tsni-rhs}
    |c1'| \approx_{|l|} |c2'|
     \qquad \text{and} \qquad
    |c2| |.->|^* |c2'|
    \ \text{.}
  \end{equation}
\end{definition}
%
In other words if we take two |l|-equivalent configurations, then for every
intermediate step taken by the first configuration, there is a corresponding
number of steps that the second configuration can take to result in a
configuration that is |l|-equivalent to the first resultant configuration.
\Red{TODO: Do we want a commutative diagram to illustrate this? Ale: I don't
  think so, it is quite standard}

We now show that our IFC language satisfies TSNI %under a particular
%scheduling policy.
%
%Specifically, we consider the language 
under the round-robin scheduler
|roundrobinf| given in Figure~\ref{fig:scheduler}.

@@

In general TSNI will not hold for an arbitrary scheduler |alpha|.
%
For example, |specLang alpha targetLang| with a scheduler that inspects a
sensitive threads current state when deciding which thread to schedule next
will in general break non-interference~\cite{Russo:Sabelfeld:CSFW06,BartheRRS07}.
%
%In Section~\toref{} we detail the preceise scheduler for which our
%specification language is TSNI.
%

However, even non-adversairal schedulers are not always safe.
Consider, for example, the sequential scheduling policy |seqf| given in
Figure~\ref{fig:scheduler}.
%
It is easy to show that |specLang seqf targetLang| does not satisfy
TSNI.
%
Concretely, consider a target language similar to |targetLangML| with an
additional expression terminal |diverge| that denotes a divergent computation,
i.e., |diverge| always reduces to |diverge| and a simple label lattice |{pub,
sec}| such that |pub canFlowTo sec|, but |sec cantFlowTo pub|.
Consider the following two configurations in this language:
\begin{code}
ic1 = iS ; fullconf 1 sec tS1 (IT(if false then diverge else true)),  fullconf 2 pub tS1 ie
ic2 = iS ; fullconf 1 sec tS1 (IT(if true then diverge else true)),   fullconf 2 pub tS2 ie
\end{code}
\Red{Ale: |when| was not defined, so I changed it to if-then-else}
These two configurations are |pub|-equivalent, but |ic1| will reduce 
(in two steps) 
to |ic1 = iS ; fullconf 2 pub tS1 (IT true)|, whereas |ic2| will not make
any progress.
%
Suppose that |ie| is a computation that writes to a |pub| channel,\footnote{
Though we do not model labeled channels, extending the calculus with such a
feature is straight forward, see Setion~\ref{sec:extensions}.}
then the |sec| tasks's decision to diverge or not is directly leaked to a
public entitiy.

% sheule: I feel like this forward reference is not useful here.
%In certain cases, however, a TSNI-safe concurrent scheduling policy such as
%|roundrobinf| cannot be considered without imposing a non-trivial burden on the
%concrete implementation.
%(Section~\ref{ref:real} describes a particular embedding with JavsScript as the
%target language.)
%
To accomodate for sequential languages, or cases where a weaker guarantee
is sufficient, we consider an alternative non-interference property called termination insensitive
non-interference (TINI).  This property can also be upheld by sequential languages at the cost
of leaking through (non)-termination~\cite{Askarov:2008}.
%
\begin{definition}[Termination insensitive non-interference (TINI)]
  A language |(C, V, .->, erasef l)| is termination
  insensitive non-interfering if for any label |l|, and configurations
  |c1, c2 member C| and |c1', c2' member V|, it holds that
  \[
    (|c1| \approx_{|l|} |c2|
    \;\;\land\;\;
    |c1| |.->|^* |c1'|
    \;\;\land\;\;
    |c2| |.->|^* |c2'|)
    \implies
    |c1'| \approx_{|l|} |c2'|
  \]
\end{definition}

@@

As shown by compilers~\cite{jif,FlowCaml}, interpreters~\cite{JSFlow}, and
libraries~\cite{Russo+:Haskell08,lio}, TINI is useful for sequential
settings. We show that our IFC language with the sequential scheduling policy
|seqf| satisfies TINI.
%
\begin{theorem}[Sequential IFC language is TINI]
  \label{thm:seq-tini}
For any target language |targetLang|, |specLang seqf targetLang| satisfies TINI.
\end{theorem}

% Local Variables:
% TeX-master: "main.tex"
% TeX-command-default: "Make"
% End:
